---
title: "cRowflow - Wine Dataset Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Wine Dataset Vignette}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(cRowflow)
library(dplyr)
library(caret)
library(plotly)
library(ggplot2)

set.seed(42)

# Create a 3D scatter plot
create_3d_plot_pca <- function(pca, color_vector, categorical=FALSE, title_addition="") {
  if (categorical) {
    pca$color <- as.factor(color_vector)
  } else {
    pca$color <- color_vector
  }
  
  plot_ly(data = data.frame(pca), 
          x = ~PC1, y = ~PC2, z = ~PC3, 
          color = ~color,
          type = "scatter3d", 
          mode = "markers") %>%
    layout(
      title = paste0("Wine Dataset PCA", title_addition),
      scene = list(
        xaxis = list(title = "PC 1"),
        yaxis = list(title = "PC 2"),
        zaxis = list(title = "PC 3")
      )
    )
}
```


```{r}
data_path <- system.file("extdata", "wine_df.csv", package = "cRowflow")
data <- read.csv(data_path)
data$X <- NULL
head(data)
```

```{r}
summary(data)
```
Features are in much different scales. We have to scale the data before proceeding to clustering.

```{r}
# Scale data.
data_scaled <- scale(data)

# Calculate PCA.
pca <- data.frame(prcomp(data_scaled)$x)
```

Let's run repeated stochastic clustering using `kmeans`. We will run the algorithm 30 times, check it's clustering stability over repeated iterations, and get the majority voting label for each sample.

Will set the number of clusters to 5 for this first experiment.
```{r}
scr <- cRowflow::stochastic_clustering_runner(
  data_scaled,
  kmeans,
  labels_name = "cluster",
  n_runs = 30,
  centers=5
)
```

Let's visualise the majority voting labels returned by `cRowflow`.
```{r}
create_3d_plot_pca(pca, scr$majority_voting_labels, T, title_addition = " - Clusters")
```

Let's now examine the stability of each element over the repeated iterations.
```{r}
summary(scr$ecc)
```
The mean ECC is 0.65 and median is 0.68. 
```{r}
create_3d_plot_pca(pca, scr$ecc, title_addition = " - ECC")
```

The most unstable points are the ones at the boundaries between clusters 2, 3, and 5.

We can use `cRowflow`'s `parameter_searcher` function to find optimal clustering parameter values that result in reproducible and robust clustering results.
```{r}
param_grid = list(centers=seq(3,8), algorithm = c("Hartigan-Wong", "Lloyd", "Forgy", "MacQueen"))

parameter_searcher <- cRowflow::parameter_searcher(
  data_scaled,
  kmeans,
  labels_name = "cluster",
  param_grid = param_grid,
  n_runs = 30,
  iter.max = 30
)
```

Let's visualise the results when changing the number of clusters and the algorithm implementation.
```{r}

cRowflow::plot_heatmap(parameter_searcher$results_df, "centers", "algorithm")
```
The heatmap displays the median ECC values for different combinations of k-means clustering algorithms and the number of cluster centers. The results indicate that lower center values (e.g., 3, 4) yield higher ECC values, with the Hartigan-Wong implementation producing more stable results in all cases. In contrast, higher center values (e.g., 8) lead to lower ECC scores across all algorithms.

Let’s further optimize clustering stability when the number of clusters (`centers`) is set to 4 and the implementation is Hartigan-Wong by identifying the optimal feature subset using a genetic algorithm. The current median ECC for that configuration is 0.79.
```{r}
set.seed(42)
genetic_fs <- cRowflow::genetic_algorithm_feature_selector(
  data_scaled,
  kmeans,
  labels_name = "cluster",
  verbose = T,
  n_generations_no_change = 5,
  centers=4,
  algorithm = "Hartigan-Wong",
  iter.max=30
)

cRowflow::plot_ga_fitness_evolution(genetic_fs$history)
```


```{r}
genetic_fs_best_fitness <- genetic_fs$best_fitness_scr_result
create_3d_plot_pca(pca, genetic_fs_best_fitness$ecc, title_addition = " - ECC AFTER GA OPTIMIZATION")
create_3d_plot_pca(pca, genetic_fs_best_fitness$majority_voting_labels, categorical = T, title_addition = " - CLUSTERS")
```

By keeping only 7 of the features we were able to produce a much more stable clustering with the same configurations (centers=4, algorithm = "Hartigan-Wong", iter.max=30). 
```{r}
colnames(data_scaled[,genetic_fs$best_features])
```
